---
title: "MUHIRE Edmond Project"
author: "MUHIRE Edmond"
date: "`r Sys.Date()`"
output: html_document
---
# THE PROJECT ASSIGNMENT OF STATISTICAL COMPUTING (SECTION 1)
# QUESTION 1: HEALTHCARE ANALYTICS
```{r}
library(foreign)
library(readxl)
library(haven)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(raster)
library(tidyverse)
```
# a) Reading and Understanding the Dataset.
# i. Read or import NHANES dataset in R.
```{r}
NHANES<-read.csv("C:/Users/USER/Desktop/Project/Dataset for R/NHANES.csv")
NHANES
View(NHANES)
```
# ii. How many variables in the dataset? What are the names of the variables in the dataset?
```{r}
# Number of variables in the data set.
Number_of_variables<-ncol(NHANES)
Number_of_variables
```
```{r}
# Names of variables in the data set.
Name_of_variables<-names(NHANES)
Name_of_variables
```
# iii. Select following column number 3, 4, 1, 7,9,10,12,14,15,16,35,40,46.
```{r}
# Selecting the columns
Select_columns<-NHANES[,c(3, 4, 1, 7,9,10,12,14,15,16,35,40,46)]
Select_columns
```
# iv. What data types are associated with each variable selected?
```{r}
str(Select_columns)
```
# v. Provide the numerical and categorical variables in the dataset
```{r}
#Numerical variables in the data set
Numerical_variables<-sapply(NHANES, function(x) is.numeric(x))
Numerical_variables_only<-NHANES[,Numerical_variables]
Numerical_variables_only
```
```{r}
#Categorical variables in the data set
categorical_variables<-sapply(NHANES, function(x) is.character(x))
categorical_variables_only<-NHANES[,categorical_variables]
categorical_variables_only
```
# vi. Provide summary statistics for numerical variables (mean, median, standard deviation)
```{r}
#Mean of numeric variables
X<-colMeans(Numerical_variables_only, na.rm=TRUE)
X
```
```{r}
#Median of numeric variables
Y<-apply(Numerical_variables_only,2,median, na.rm=TRUE)
Y
#The means and the medians are almost equal so the data are normal distributed.
```
```{r}
#Standard deviation of numeric variables
Z<-apply(Numerical_variables_only,2,sd,na.rm=TRUE)
Z
```
# vii. Are there any missing values and duplicates in the dataset, and if so, how many?
```{r}
#Showing if there are the missing values
any_missing_value<-any(is.na(NHANES))
any_missing_value
if (any_missing_value){print("There are the missing values")} else {print("There are no missing values")}
```
```{r}
# Number of missing values
Number_of_missing_values<-sum(is.na(NHANES))
Number_of_missing_values
```
```{r}
#Showing if there are the duplicates in data set
Any_duplicate<-any(duplicated(NHANES))
Any_duplicate
if (Any_duplicate){print("There are duplicates")} else {print("There are no dupilcates")}
```
```{r}
#Number of duplicates in the data set
Number_of_duplicates<-sum(duplicated(NHANES))
Number_of_duplicates
```
# viii. Visualize outliers in Age using boxplot. (remember to set limit if your limits are too large)
```{r}
boxplot(NHANES$Age, main = "Boxplot of Age", ylab = "Age",ylim = c(0, 110),col = "green")
```
#b) Data Cleaning, manipulation and Exploratory Data Analysis.
#i. Fill the missing values in the ' TotChol' column by zero (0) and check again if there is no missing in that column
```{r}
#Filling the missing values in the "Totchol" column by zero (0)
NHANES$TotChol[is.na(NHANES$TotChol)]<-0
NHANES$TotChol[is.na(NHANES$TotChol)]
```
```{r}
#Checking again if there is no missing in the column
any_missing_after_filling<-any(is.na(NHANES$TotChol))
any_missing_after_filling
#Using if else condition to show there is no missing value
if (any_missing_after_filling) {print("There is missing vale")} else {print ("There is no missing values")}
```
# ii. Remove all duplicated identified if any and check again if no duplicates
```{r}
#Showing if there the duplicates
any_duplicates_before_removal<-any(duplicated(NHANES))
any_duplicates_before_removal

#Using the if else condition to show if duplicates are present
if (any_duplicates_before_removal) {print("There are the duplicates")} else {print("There are no duplicates")}
```
